Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-05-17 WRITEUP/ms.tex
section 6 commented

In this section, we compare the observable quantities of MCCs obtained from the simulations (e.g., $\sigma_{\rm gas}$, $R$, $M_{\rm gas}$,
$\Sigma_{\rm gas}$) with those of molecular gas structures from \obs at.
Excluding the biggest MCC identified in each evolutionary stage (i.e., the top right cluster of points in the
linewidth-size relation of \Fig{larsons_single} and the top left panel of \Fig{alpha16-28}), which correspond to
the main disk of \flower, the MCCs of \flower have dynamics largely similar to those observed in
$z$\ssim2 spatially resolved studies of gas-rich star-forming
galaxies, in terms of their velocity dispersions, sizes, and gas
surface densities (\Fig{larsons_single}; see e.g.,
\citealt{Swinbank11a}), with sizes of the order of $R\simeq$\,100\,pc
and velocity dispersions of $\sigma_{\rm gas}\simeq$\,20$-$80\,\kms.
Their velocity dispersions are also comparable to those observed in the inner
Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a}), which lies along the
locus of $\sigma_{\rm gas}\propto R^{0.56}$.
Such high velocity dispersions and surface densities can be explained partly because
these MCCs are located in the nuclear regions of the galaxy, where the potential well is the deepest,
and also from the injection of kinetic energy from recent
\SF as \flower is assembling its stellar mass (given the high Mach numbers of MCCs; \Sec{distribution}).
In fact, by the accreting stage (see \Fig{SFH}a) at $z\simeq$\,7.2,
\flower has assembled a stellar mass of
$M_\star$\eq7.5\E{9}\,\Msun.
Furthermore, a comparison of the velocity dispersion from bulk motions versus that from turbulence (\Fig{vv}) suggests that
the high total velocity dispersion seen in the MCCs of \flower is partly a result of large-scale rotation, which is even more important in
its later stage of evolution as \flower is settling into a rotating disk galaxy. The implication of this in
interpreting observational results is further discussed in the next subsection (\Sec{rot}).

The high pressure observed in \flower (see \citealt{Pallottini17a})
is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
For instance, in local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}.
% Tasker09a
On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{kohandel:2019} and higher velocity accretion/SN-driven outflows \citep{gallerani:2018}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g., \citealt{Swinbank11a}) --- but the observed size scales could also be biased high due to the resolution limit.

@@

The local sound speed of all identified molecular complexes is typically much smaller than their non-thermal (turbulent) velocities,
as non-thermal pressure dominates thermal pressure for dense gas \citep{Pallottini17b}, i.e. $c_s^2 \ll \sigma_{\rm NT}^2$.
 In particular, the Mach number averaged on the the MCCs distribution at the highest density threshold is $\mathcal{M} \simeq6$; this is consistent with the analysis done in \citep{Vallini18a}, that finds a global Mach number of $\mathcal{M} \sim 10$ for \flower.
% allss: 0.32
% In [13]: print(get_Mach_all_clouds(ss11))
%[10.62623949  3.10713402  7.0525831   3.336829   10.36052827  7.24157821
%  7.12323137  6.75282955  9.94086767  7.47769907  3.39894958  3.27304956
%  6.69290803 11.44844494  7.03862716  3.16574077  5.1015787   5.67749554
%  7.35946615  6.89475134  7.2433421  10.57019778 11.78414139  3.93499799
%  6.86173652  7.39686112  8.37197686  6.71749578  2.65304172  6.49617895
%  5.15095491  5.6606244   2.98923849  3.65475035  7.37703114  4.5669722
%  7.18720926  7.35017082  6.03000248  6.36373235 10.16701766  8.3894084
%  3.09284034  4.37994321  4.46497953  5.264101    8.11686     3.05183157
%  7.10365783  6.09513017  8.72517656  5.59659665  3.95290202  9.22726672
%  4.94960657  6.19757621  4.63335759  3.32670022  3.46706217  5.58358069
%  3.41808098  8.31386962  4.26357159  3.684986    4.01853538  3.62917357
%  6.05586465  9.82127908  6.68366736 11.0791874  13.28965621 10.42021747
%  3.04409362  7.21509021  3.63699627  8.12503492  3.20158298]
% allss: 18.96
%In [15]: print(get_Mach_all_clouds(ss12))
%[15.25049689  6.74580503  6.98246499  5.17070655  6.470997    7.02626558
%  4.42663778  6.34460542  4.90603266  3.66900884  4.03260414  6.9539893
%  8.36247744  4.89164338  4.12695121  6.55187227  6.15391173  3.56580625
%  5.74818066  3.81067175  6.65393948  9.10100946  6.35039889  4.78667156
%  5.05517168  6.93886988  5.4830306 ]
% see also e.g., test_brute/leaf_fields_ss16-28/18.96_10_fields/Mach_cloud-mass.png
% This is consistent with the global analysis done on \flower by \citet{Vallini18a}.

We show the gas velocity dispersions from bulk motions versus turbulent motions for MCCs in the accreting
and quiescent phase in \Fig{vv}. We note that \flower displays a disturbed morphology in the accreting phase
but a disk-like morphology in the quiescent phase
(see \Fig{MCC} and \Fig{MCC28}).
Most MCCs in the accreting phase have turbulent velocity dispersions higher than the bulk motion velocity dispersions,
whereas the opposite is true in the quiescent phase. 
In addition, the two velocity dispersion terms are of the same order of magnitude. % as the two are connected via the Euler equation
{\em This indicates that gas velocity dispersions observed on similar physical scales and resolutions as presented here
have an important contribution from large-scale rotation.} Hence, the linewidths observed in \highz molecular gas structures
on these scales should {\em not} be taken as a direct evidence they are intrinsically more turbulent than those in the local ISM,
and the comparison and interpretation of the dynamics of gas structures with local clouds using e.g., the Larson's relations is non-trivial.
\citet{Behrendt16a} draw a similar conclusion using pc-scale resolution simulation: 
large-scale bulk motions may mimic turbulent velocity dispersion due to  
beam smearing, even though contributions due to self-rotation of each individual sub-MCCs is washed out within the beam.

@@

\begin{figure*}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.92\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_2.0}.pdf}
\includegraphics[trim=0 0 0 65, clip, width=0.95\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_0.8}.pdf}
\caption{
Toomre $Q$ maps derived from the
central $r$\eq2\,kpc (top row) and $r$\eq0.8\,kpc (bottom row) of \flower.
Gas-only $Q_{\rm gas}$ is shown in the left panels and stellar-only $Q_\star$ is shown in the middle panels.
Maps of the effective two-component Toomre $Q_{\rm eff}$ parameter are shown in the right panels.
All maps are projected onto the $xy$-plane.
Positions of MCCs identified with $n_{\rm cut}$\eq6.81\,cm$^{-3}$ are overplotted as star symbols.
A smoothing length of 30\,pc has been applied to the maps.
A divergent colormap is used for the Toomre $Q$ maps to facilitate
identification of regions above and below $\log{Q}$\eq0.
Some MCCs lie in regions of $\log{Q_{\rm eff}}\gtrsim0$, where regions of $\log{Q_{\rm eff}}\lesssim0$
are likely gravitationally unstable.
Close resemblance of the $Q_\star$ and $Q_{\rm eff}$ maps in the {\em central region} of \flower
indicates that the stellar component plays an important role in
governing the stability of the MCCs against $m=0$  perturbations,
highlighting the importance of accounting for stellar contribution when examining the stability of molecular gas in relatively evolved and enriched systems at high redshifts.
\label{fig:Qeff}}
\end{figure*}

Maps of the Toomre $Q$ parameters for the gas, stars, and the effective two-component $Q_{\rm eff}$ are shown in \Fig{Qeff}
(see also \Fig{h2density} for example of MCCs identified in this evolutionary stage of \flower
using different $n_{\rm cut}$).

Close resemblance of the $Q_{\rm eff}$ and $Q_{\star}$ maps indicates that contributions from the stellar component play an important role in governing the stability of the MCCs against $m$\eq0 perturbations. This can be understood since stars in \flower dominates the central part of the galaxy in mass, so their gravitational potential provides a non-negligible contribution to the instability.
Similarly, contribution from the thickness of the disk is important in \flower since its disk is warped and
has a scale height-to-radius ratio of $h/r_{\rm gal}$\ssim150\,pc/1\,kpc\,$\simeq$\,0.15.

That is, some MCCs are found in regions of $Q_{\rm gas}$\ssim1, which is consistent with the expectation that
they correspond to regions of high surface densities that are gravitationally unstable.
Note, however, that when including the stabilizing effects due to the stellar potential of \flower and the
thickness of its disk (via $\sigma_z$, see Equation~\ref{eqn:q_eff}), some of these MCCs are
consistent with $Q_{\rm eff} > 1$.
This demonstrates the importance of accounting for stellar contribution and disk thickness
when examining the stability of molecular gas structures.
This consideration is especially relevant for the relatively evolved and enriched systems at high redshift,
that are preferentially being imaged at high resolution with ALMA.
In the outer regions of \flower, on the other hand, $Q_{\rm eff}$ resembles $Q_{\rm gas}$, with
both $\log{Q_{\rm gas}}$ and $\log{Q_{\rm eff}} < -1$.
Notably, these are the more gas-rich regions (see \Fig{sigma}).

The large virial parameters seen in some MCCs of \flower can be understood
by first noting that fragmentation {\em can} happen in regions of low $Q$
(if we only consider instability against axisymmetric perturbations),
but further evolution/collapse depends on the dynamics or equation of state of the gas.
%
Such fragmentation is expected to take place at the critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$.
Second, this fragmentation scale is greater than the typical size of GMCs.
%
This could be interpreted as a result of instability setting the scales for fragmentation,
{\em but the truly star-forming regions correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales}.
%
Thus, the high virial parameter found for some MCCs in \flower
indicates that they are not the collapsing structures (but the virial parameter varies depending on the evolutionary stage
of \flower) and are found in regions of $Q_{\rm eff}>1$.
On the other hand, MCCs in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable
and \SF is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales,
where energy quickly dissipates. % with low virial parameters.

If most of the MCCs in \flower are stable against collapse,
how does it sustain its high SFR of $\sim$70\,\Msun\,yr\pmOne in the starburst phase?
Such tension in the interpretation is understood as follows.
%
First, our results are limited by the resolution of the simulation, i.e., in \flower~we do not resolve \SF in {\it cores} and {\it clumps} as in the classical molecular cloud hierarchy. This re-iterates the statement about low $Q$ above.
%
Second, due to turbulence dissipation on small-scales, such
(sub-)structures are no longer supported by differential rotation.
This in turn enables sub-regions of the MCCs to collapse and potentially form
OB associations out of these unbound MCCs \citep{Clark04a, Clark05a}.
% Mechanisms causing non-axisymmetric perturbations (e.g.,
% arms) to the gravitational potential \MM{Gravitational instability is
%   that mechanism (see e.g.\ Wada et al.'s classic review of the
%   subject (2011, doi:10.1017/S1743921311000640), so this isn't really a separate mechanism.}
% can induce orbit crossing, shocks and dissipation in gas, promoting turbulence dissipation and collapse on smaller scales to trigger \SF. These mechanisms are likely important given high $Q_{\rm eff}$ found across \flower and its few tens of solar masses per year SFR, albeit sporadic.
Further, the SFR of \flower is very bursty and high SFR stages are only transient.

>>>>

In this section, we compare the observable quantities of MCCs obtained from the simulations (e.g., $\sigma_{\rm gas}$, $R$, $M_{\rm gas}$, $\Sigma_{\rm gas}$) with those of molecular gas structures from \obs at lower redshift.
%
We exclude for the moment the biggest MCC identified in each evolutionary stage, which correspond to the main disk of \flower (i.e., the top right cluster of points in the linewidth-size relation of \Fig{larsons_single} and the top left panel of \Fig{alpha16-28}).

The MCCs of \flower have dynamics largely similar to those observed in $z$\ssim2 spatially resolved studies of gas-rich star-forming galaxies, in terms of their velocity dispersions, sizes, and gas surface densities (\Fig{larsons_single}; see e.g., \citealt{Swinbank11a}), with sizes of the order of $R\simeq$\,100\,pc and velocity dispersions of $\sigma_{\rm gas}\simeq$\,20$-$80\,\kms.
%
Their velocity dispersions are also comparable to those observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a}), which lies along the locus of $\sigma_{\rm gas}\propto R^{0.56}$. Such high velocity dispersions and surface densities can be explained partly because these MCCs are located in the nuclear regions of the galaxy, where the potential well is the deepest, and also from the injection of kinetic energy from recent \SF as \flower is assembling its stellar mass (given the high Mach numbers of MCCs; \Sec{distribution}). In fact, by the accreting stage (see \Fig{SFH}a) at $z\simeq$\,7.2, \flower has assembled a stellar mass of $M_\star$\eq7.5\E{9}\,\Msun.
%
\AP{I do not fully understand the following} Furthermore, the comparison of the velocity dispersion from bulk motions versus that from turbulence (\Fig{vv}) suggests that the high total velocity dispersion seen in the MCCs of \flower is partly a result of large-scale rotation, which is even more important in its later stage of evolution as \flower is settling into a rotating disk galaxy. The implication of this in interpreting observational results is further discussed in the next subsection (\Sec{rot}).

The high pressure observed in \flower (see \citealt{Pallottini17a}) is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{kohandel:2019} and higher velocity accretion/SN-driven outflows \citep{gallerani:2018}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g., \citealt{Swinbank11a}) --- but the observed size scales could also be biased high due to the resolution limit.

@@

The local sound speed of all identified molecular complexes is typically much smaller than their non-thermal (turbulent) velocities,
as non-thermal pressure dominates thermal pressure for dense gas \citep{Pallottini17b}, i.e. $c_s^2 \ll \sigma_{\rm NT}^2$. In particular, the Mach number averaged on the the MCCs distribution at the highest density threshold is $\mathcal{M} \simeq6$; this is consistent with the analysis done in \citep{Vallini18a}, that finds a global Mach number of $\mathcal{M} \sim 10$ for \flower.
%
This can be further analyzed as follows. In \Fig{vv} we show the gas velocity dispersions from bulk motions versus turbulent motions for MCCs in the accreting and quiescent phase. Recall that \flower displays a disturbed morphology in the accreting phase but a disk-like morphology in the quiescent phase (see \Fig{MCC} and \Fig{MCC28}).
%
Most MCCs in the accreting phase have turbulent velocity dispersions higher than the bulk motion velocity dispersions, whereas the opposite is true in the quiescent phase. In addition, the two velocity dispersion terms are of the same order of magnitude.
% as the two are connected via the Euler equation
\AP{I still do not get the point; bulk motion is a dispersion, i.e. should not receive contribution from the rotation of the system; a possibility might be to calulate the specific angular momentum ($\vec{x}\wedge\vec{v}$ averaged using \Eq{defineaverage}) of the system and compare it to the one for gas in circular motion} {\em This indicates that gas velocity dispersions observed on similar physical scales and resolutions as presented here have an important contribution from large-scale rotation.}
%
Hence, the linewidths observed in \highz molecular gas structures on these scales should {\em not} be taken as a direct evidence they are intrinsically more turbulent than those in the local ISM, and the comparison and interpretation of the dynamics of gas structures with local clouds using e.g., the Larson's relations is non-trivial. \citet{Behrendt16a} draw a similar conclusion using pc-scale resolution simulation: large-scale bulk motions may mimic turbulent velocity dispersion due to beam smearing, even though contributions due to self-rotation of each individual sub-MCCs is washed out within the beam \AP{they argue that self-rotation is not important, but i do not see how we match the result with our analysis}.

\AP{I would not remove the mass-size relation discussion (commented now in the text)}

@@

\begin{figure*}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.92\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_2.0}.pdf}
\includegraphics[trim=0 0 0 65, clip, width=0.95\textwidth]{\figpath/{ss16_toomre_combined_3by1_0_0.8}.pdf}
\caption{
Toomre $Q$ maps derived from the
central $r$\eq2\,kpc (top row) and $r$\eq0.8\,kpc (bottom row) of \flower.
Gas-only $Q_{\rm gas}$ is shown in the left panels and stellar-only $Q_\star$ is shown in the middle panels.
Maps of the effective two-component Toomre $Q_{\rm eff}$ parameter are shown in the right panels.
All maps are projected onto the $xy$-plane.
Positions of MCCs identified with $n_{\rm cut}$\eq6.81\,cm$^{-3}$ are overplotted as star symbols.
A smoothing length of 30\,pc has been applied to the maps.
A divergent colormap is used for the Toomre $Q$ maps to facilitate
identification of regions above and below $\log{Q}$\eq0.
Some MCCs lie in regions of $\log{Q_{\rm eff}}\gtrsim0$, where regions of $\log{Q_{\rm eff}}\lesssim0$
are likely gravitationally unstable.
Close resemblance of the $Q_\star$ and $Q_{\rm eff}$ maps in the {\em central region} of \flower
indicates that the stellar component plays an important role in
governing the stability of the MCCs against $m=0$  perturbations,
highlighting the importance of accounting for stellar contribution when examining the stability of molecular gas in relatively evolved and enriched systems at high redshifts. \AP{maybe it is better to put a single set of maps, with FOV from -1.5 to 1.5 kpc}
\label{fig:Qeff}}
\end{figure*}

\AP{Q was introduced at page 3, but then not used. I think some intro sentence should be added here}
Maps of the Toomre $Q$ parameters for the gas, stars, and the effective two-component $Q_{\rm eff}$ are shown in \Fig{Qeff} (see also \Fig{h2density} for example of MCCs identified in this evolutionary stage of \flower using different $n_{\rm cut}$).

Close resemblance of the $Q_{\rm eff}$ and $Q_{\star}$ maps indicates that contributions from the stellar component play an important role in governing the stability of the MCCs against \AP{should we introduce this classification of the perturbations in the \Sec{Q}} $m$\eq0 perturbations. This can be understood since stars in \flower dominates the central part of the galaxy in mass, so their gravitational potential provides a non-negligible contribution to the instability. Similarly, contribution from the thickness of the disk is important in \flower since its disk is warped and has a scale height-to-radius ratio of $h/r_{\rm gal}$\ssim150\,pc/1\,kpc\,$\simeq$\,0.15.

That is, some MCCs are found in regions of $Q_{\rm gas}$\ssim1, which is consistent with the expectation that they correspond to regions of high surface densities that are gravitationally unstable. Note, however, that when including the stabilizing effects due to the stellar potential of \flower and the thickness of its disk (via $\sigma_z$, see Equation~\ref{eqn:q_eff}), some of these MCCs are consistent with $Q_{\rm eff} > 1$.
%
This demonstrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures. This consideration is especially relevant for the relatively evolved and enriched systems at high redshift, that are preferentially being imaged at high resolution with ALMA. In the outer regions of \flower, on the other hand, $Q_{\rm eff}$ resembles $Q_{\rm gas}$, with both $\log{Q_{\rm gas}}$ and $\log{Q_{\rm eff}} < -1$. Notably, these are the more gas-rich regions (see \Fig{sigma}).

The large virial parameters seen in some MCCs of \flower can be understood by first noting that fragmentation {\em can} happen in regions of low $Q$ (if we only consider instability against axisymmetric perturbations), but further evolution/collapse depends on the dynamics or equation of state of the gas.
%
Such fragmentation is expected to take place at the critical scale length $\lambda_{\rm crit} < 2 \pi^2 G \Sigma / \kappa^2$. Second, this fragmentation scale is greater than the typical size of GMCs.
%
This could be interpreted as a result of instability setting the scales for fragmentation, {\em but the truly star-forming regions correspond to the collapsing, denser, and cooler molecular structures that are on smaller scales}.
%
Thus, \AP{from the sentence, it is unclear if the Q analysis is consistent or is in an independent indication with respect to the virial parameter analysis} the high virial parameter found for some MCCs in \flower indicates that they are not the collapsing structures (but the virial parameter varies depending on the evolutionary stage of \flower) and are found in regions of $Q_{\rm eff}>1$. On the other hand, MCCs in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable and \SF is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, where energy quickly dissipates.

If most of the MCCs in \flower are stable against collapse, how does it sustain its high SFR of $\sim$70\,\Msun\,yr\pmOne in the starburst phase? Such tension in the interpretation is understood as follows.
%
First, due to turbulence dissipation on small-scales, such (sub-)structures are no longer supported by differential rotation. This in turn enables sub-regions of the MCCs to collapse and potentially form OB associations out of these unbound MCCs \citep{Clark04a, Clark05a}.
% Mechanisms causing non-axisymmetric perturbations (e.g., arms) to the gravitational potential \MM{Gravitational instability is that mechanism (see e.g.\ Wada et al.'s classic review of the subject (2011, doi:10.1017/S1743921311000640), so this isn't really a separate mechanism.} can induce orbit crossing, shocks and dissipation in gas, promoting turbulence dissipation and collapse on smaller scales to trigger \SF. These mechanisms are likely important given high $Q_{\rm eff}$ found across \flower and its few tens of solar masses per year SFR, albeit sporadic.
Further, the SFR of \flower is very bursty and high SFR stages are only transient.
%
Finally, our results are limited by the resolution of the simulation, i.e., in \flower~we do not resolve \SF in {\it cores} and {\it clumps} as in the classical molecular cloud hierarchy. This re-iterates the statement about low $Q$ above.
